\plaintitle

### 2017-01-06 drafts/labor/piecework_lit.tex
lit review not finished but ready for more feedback; switching to case studies to hack stuff up

\ali{I'm interleaving the paragraphs, so this paragraph needs to be refactored a bit to reflect that instead of suggesting that the road map will look like two distinct threads.
I'll probably try to take a shot at it in a bit, but for now don't take the wording too seriously.}

@@

\ali{
So I'm thinking that this should flow like this:
``We'll try to understand what piecework was by tracing the development of and discourse around piecework.''
This'll give us the ability to switch back and forth between the timeline of the practice of piecework
--- that is, the ``piecework primer'' stuff down below ---
and the contemporary/concurrent research going on around the same time
(i.e. this stuff immediately following).

I'm going to interleave the paragraphs, but
if you think that's a flawed approach, then I'll bet you won't like what you'll see below that much.%
}

\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over the years,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions,
tracing an outline of % a working understanding of
\textit{what piecework is}
alongside piecework's contemporary developments in practice.
\ali{This is worded a little awkwardly;
I want to say that we're looking at
the discourse about piecework at the same time that we look at
how piecework is unfolding. (wait, is that how I should word it?)}

@@

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
\ali{I'm not sure; I'll look for some data.}

% \ali{\textbackslash topic'ing everything really makes clear when I have a run--on sentence\dots}
\topic{By \citeyear{hughRaynbirdTaskWork} we find % \citeauthor{hughRaynbirdTaskWork} offers
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay % 's \citetitle{hughRaynbirdTaskWork}
(where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
particularly driven toward encapsulating the manual labor of farmwork.}
\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'':
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers a more intuitive sense;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

@@

This frustrations of poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

@@

% \msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
Between 1938 and 1942, piecework in a number of metal workers had climbed as steeply as from 11\% to 60\%~\cite{hart2005piecework};
by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions piecework
--- where more than 50\% worked under incentive wage plans ---
were principally in clothes--making (e.g. hosiery, footwear, and garments).
% \msb{How quickly did it plummet? By when was it gone? Where is it now?}
% \msb{This sentence is empty calorie:}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.

\ali{let's try cutting it\dots
\msb{Very few calorie sentence:}
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance
\item the shifting of these industries (e.g., manufacturing, clothing)
to other countries
\item the quality of ``multidimensional'' work becoming too difficult to evaluate~\cite{hart2013rise}.
\end{inlinelist}}

\ali{This feels like the Australia of paragraphs: one sentence constitutes the whole paragraph.
Should I add some analysis here?}
In summary, piecework:
\begin{inlinelist}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models
  \item afforded workers freedom in when and how much to work
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. 
\msb{This transition seems misplaced with the Piecework primer and Case Studies subsections up next:}
In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.

% \onlyinsubfile{\clearpage}
% \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
% \msb{This section is so brief that it feels like a bit of a cheat. I think we need to frame it differently. Maybe that it should survey the different ``eras'' of piecework? And each paragraph is laying out a major historical era that we will return to later? Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
% \ali{Do you mean that we'd have (sub)subsections under this? I'm tempted to merge this subsection into the ``\namerefl{sec:whatIsPiecework}'' subsection instead. Not exactly sure how that would work (i.e. would I interleave the history or just append it, basically removing the header).}
\ali{there's no need for this paragraph:

\topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
This section will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.}
% While the previous section provided a \textit{definition} of piecework,
% this section attempts to shine a light on the \textit{zeitgeist} of piecework;
% as close to a \textit{thick description} of piecework
% as we can afford~\cite{geertz1973interpretation}.
% \msb{is this really thick? as a reader it flies by really quickly and super high level...}

% \msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
% \citeauthor{hart2013rise} point out that the Second World War,
% which called millions of Americans to military service,
% necessitated the rapid training and employment of
% a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.

% \msb{The topic sentence is about management, but the content is about discontent. Which is it?}

% \onlyinsubfile{\clearpage}
\subsection{Case studies in piecework}
Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
first, railroad and other industrial workers;
second, Airy's employment of \textit{human computers};
and third, domestic and farm work (in particular, the ``match--girls'').
Here, we introduce the basic historical details of these three cases.
We begin with the most familiar case (industrial workers),
working backwards through the \nth{20} and then \nth{19} century,
tracing the origins of the ideas that \dots
% \msb{incomplete sentence}

@@

% \ali{maybe go backwards on this}

% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
% \onlyinsubfile{\clearpage}
\subsubsection{Railroad and other Industrial Workers}
{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
It was here that
scientific management, Fordism, and Taylorism
dramatically influenced how workers were managed and
the ways in which workers were perceived and envisioned paradigmatically;
it's here we'll start our overview of piecework's history through case studies. \msb{we're not starting an overview of history; aren't we midway through it?}

@@

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
\msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
\msb{Why is Taylor relevant to this case study?}
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

% \onlyinsubfile{\clearpage}
\subsubsection{Airy's Computers}
\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
% \ali{also talk about the challenge/competition Airy was in with other astronomers,
% and the approach of verifying worker results}
% Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.
% For one thing,
Airy also instituted a policy of firing his computers once they reached age 23
\ali{this has a ``gosh isn't that interesting?'' feeling to it.
should i just leave it out until later when there's something to say about it
(namely, that the mode of work was initially designed to stymie professional growth)?}.

\ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
First, it eliminated any potential to advance professionally, as
workers' careers in this area ended relatively early in their careers,
and without formal education in mathematics they struggled to find work for which their experience was meaningful.
Second, it limited workers' ability to organize
by ensuring that workers hardly spent sufficient time to successfully rally their peers.
}

>>>>

@@

\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over the years,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions,
tracing an outline of % a working understanding of
\textit{what piecework is}
alongside piecework's contemporary developments in practice.

@@

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
\ali{I'm not sure; I'll look for some data.

update: hard to find anything saying ``piecework represented eleventy billion dollars of the U.S. economy'', which is what I'd like to say,
but I can say something like ``at its height, $k\%$ of garment workers were pieceworkers'' or something.
that seems a bit wishy--washy though; I'd like to say something like
``$k\%$ of the \textit{entire U.S. workforce} was working under some sort of piecework regime''.}

% \ali{\textbackslash topic'ing everything really makes clear when I have a run--on sentence\dots}
\topic{By \citeyear{hughRaynbirdTaskWork} we find % \citeauthor{hughRaynbirdTaskWork} offers
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay % 's \citetitle{hughRaynbirdTaskWork}
(where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
particularly driven toward encapsulating the manual labor of farmwork.}
\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'' ---
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers a more intuitive sense;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

@@

@@

% \msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
Between 1938 and 1942, piecework in a number of metal workers had climbed as steeply as from 11\% to 60\%~\cite{hart2005piecework};
by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions piecework
--- where more than 50\% worked under incentive wage plans ---
were principally in clothes--making (e.g. hosiery, footwear, and garments).
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.

\ali{``, ultimately arguing that\dots''?}

In summary, piecework:
\begin{inlinelist}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models
  \item afforded workers freedom in when and how much to work
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well.

\subsection{Case studies in piecework}
\topic{Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
railroad and other industrial workers;
Airy's employment of \textit{human computers};
and domestic and farm work (in particular, the ``match--girls'').}
We'll introduce the facts of these three cases,
beginning with the most familiar case of piecework
--- industrial work, such as assembly line manufacturing ---
working backwards through the \nth{20} and \nth{19} century.
In doing so, we'll trace the history of piecework
while also framing the later analysis of the major research threads we named earlier
(questions toward complexity, decomposition, and relationships, respectively).

@@

\subsubsection{Railroad and other Industrial Workers}
\topic{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
It was here that
scientific management, Fordism, and Taylorism
dramatically influenced how workers were managed and
the ways in which workers were perceived and envisioned paradigmatically;
it's here we'll start our overview of piecework's history through case studies. \msb{we're not starting an overview of history; aren't we midway through it?}

@@

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
\msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
\msb{Why is Taylor relevant to this case study?}
\ali{I'm just clumsily interleaving the contemporary literature with the events.
I'd like to say ``This is the geist of work in their time'';
is this a bad idea?}
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

% \onlyinsubfile{\clearpage}
\subsubsection{Airy's Computers}
\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
Airy also instituted a policy of firing his computers once they reached age 23
\ali{this has a ``gosh isn't that interesting?'' feeling to it.
should i just leave it out until later when there's something to say about it
(namely, that the mode of work was initially designed to stymie professional growth)?}%
.

\ali{This should go later someplace, but not clear where:

This practice ensured two outcomes that arguably disfavored workers.
First, it eliminated any potential to advance professionally, as
workers' careers in this area ended relatively early in their careers,
and without formal education in mathematics they struggled to find work for which their experience was meaningful.
Second, it limited workers' ability to organize
by ensuring that workers hardly spent sufficient time to successfully rally their peers.}
